---
title             : "What might books be teaching young children about gender?"
shorttitle        : "Gender in children's books"

author:  
  - name          : "XXX"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "XXX"
    email         : "XXX"
  - name          : "XX"
    affiliation   : "2"
    email         : "cooperborken@wisc.edu"
  - name          : "XXX"
    affiliation   : "2"
    email         : "econverse@wisc.edu"
  - name          : "XXX"
    affiliation   : "2"
    email         : "glupyan@wisc.edu"
  - name          : "XXX"
    affiliation   : "2"
    email         : "seidenberg@wisc.edu "
affiliation:
  - id            : "1"
    institution   : "XXX"
  - id            : "2"
    institution   : "XXX"
    
author_note: | 
abstract: |
  We investigated how gender is represented in children's books using a 200,000 word corpus comprising 247 popular, contemporary books for young children (0-5 years). Using human judgments and word co-occurrence data, we quantified the gender biases of words within the corpus and within individual books. We find that children's books contain large numbers of words that adults judge gendered. Semantic analyses based on co-occurrence data yielded word clusters related to gender stereotypes (e.g., feminine: emotions; masculine: tools). Co-occurrence data also indicate that books instantiate gender stereotypes found in other research (e.g., girls are better at reading and boys at math). Finally, we used large-scale data to estimate the gender distribution of the audience for individual books, and find that children tend to be exposed to gender stereotypes for their own gender. Together the data suggest that children's books may be an early source of gender stereotypes.
  
  STATMENT OF RELEVANCE:  Gender stereotypes such as that girls are better at reading and boys are better at math have origins in early childhood. We examined the extent to which popular children's books may expose young 0-5 year old children to information about gender stereotypes. Our data suggest that children's books have rich information about gender stereotypes, and that some of these stereotypes are more strongly represented in children's books than in adult fiction. These findings suggest that popular children's books may be unintentionally teaching young children about gender stereotypes.
  
keywords          : "reading, gender, language development"
wordcount         : "1972 (excluding methods and results)"
bibliography      : ["references.bib"]
header-includes:
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{graphicx}
#  - \usepackage{booktabs}
  - \usepackage{tabu}
  - \usepackage{makecell}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \usepackage{pbox}
  - \usepackage{hyphsubst}
  - \floatplacement{figure}{t!} # make every figure with caption = t
  - \raggedbottom
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\textfraction}{.1}
  - \renewcommand{\floatpagefraction}{1}


figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
documentclass     : "apa6"
lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
  

---

```{r load_utility_packages, include = FALSE}
library(papaja)
library(rmarkdown)
library(tidyverse) 
library(here)
library(broom)
library(janitor)
library(kableExtra)
library(tidyboot)
library(ggrepel)
library(ggpmisc)
library(langcog)
library(glue)
library(esc)
library(compute.es)
source(here("writeups/paper/paper_helpers.R"))
```

```{r global_options, include = FALSE}
#redoc::redoc(highlight_outputs = FALSE)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      cache = FALSE,
                      echo = FALSE,
                      out.extra = "",
                      fig.pos = "t!",  sanitize = TRUE)
```
Many gender stereotypes held by adults have origins in early childhood. Beginning in their second year, children already exhibit knowledge of behaviors that are stereotypically feminine (e.g., vacuuming), masculine (e.g., building), and neutral [e.g., sleeping; @poulin2002men]. By age three, children distinguish individuals by gender, race, and age [@shutts2010social]. By age five, children have developed “a constellation of stereotypes about gender (often amusing and incorrect) that they apply to themselves and others” [@martin2004children]. For example, preschoolers act in accordance with the stereotype that girls are better at reading while boys are better at math [@cvencek2011math], and that girls are less likely than boys to be “very, very smart” [@bian2017gender]. 

The _sources_ of this knowledge are less well understood. Certainly, some of what children know about gender characteristics comes from their own and observed interactions with adults [@hilliard2010differing]. Toys and activities are often gender stereotyped in home, day care, and preschool social settings [@weisgram2014pink]. Another source of information is language. Children commonly receive feedback from adults about gender-normative activities, e.g., girls more often receive adult linguistic feedback for dress and helping behaviors, whereas boys receive comments on their size and physical skills [@chick2002impact], and children are surprisingly sensitive to seemingly small differences in linguistic descriptions of gender-related information. For example, @cimpian2011generic found that when a novel game was introduced to children using a generic subject (“Girls are really good at a game called 'gorp'") they were more likely to associate it with a gender than when the game was introduced with a specific subject (“There is a girl who is good at...”). This sensitivity may in part arise from an essentialist bias---a tendency to treat categories such as male and female as distinct with respect to visible, inferred, and assumed characteristics [@gelman2000gender]. 

Here, we examine a particular potentially pervasive source of information about gender---books directed at young children. The practice of reading to young children (also called “shared reading”) has been widely encouraged because of its numerous benefits [@bus1995joint; @duursma2008reading; @high2014literacy]. Shared reading marks the child’s entré to literacy and facilitates its development [@snow1998preventing]. Unlike everyday speech to children, linguistic input from books exposes children to information beyond their immediate experience, and therefore may be an especially powerful way of conveying gender stereotypes. 

Much of the past work on how gender is represented in books has used “content analysis” which emphasizes detailed analyses of a small number of texts. For example, @diekman2004learning presented college students with books for middle-schoolers previously categorized as “sexist” or “nonsexist”. Students rated them on a 72-item questionnaire. Questions probed whether books conveyed common gender stereotypes and inequalities, such as “Males, but not females, are shown as dominant” and “The book depicts female characters as the natural servants of male characters.” The results suggested that gender differences and inequalities were expressed even in books intended to be nonsexist.

Our goal was to conduct a broader analysis of gender representation in books aimed at young (0-5 year old) children and to better understand who is being exposed to them. We operationalize the notion of gender representations by examining the extent to which individual concepts are differentially associated with a particular gender. By using human norms and automated analyses of distributional semantics, our approach provides a scalable and reproducible method of estimating gender biases without requiring explicit judgments of pre-specified properties of texts [as in studies such as @diekman2004learning].

We begin by describing the creation of the Wisconsin Children’s Book Corpus (WCBC). We then assess the genderedness of words used in these books. We measure genderedness using adult judgments and describe how word gender bias relates to other word properties (e.g., age of acquisition; concreteness; Study 1A). We next characterize the semantics of word gender biases using word embedding models, and compare the word biases in our corpus to those in adult fiction (Study 1B). In Study 2, we quantify the gender bias of individual books. In Study 3, we use automated analyses of book reviews to estimate whether each book is being read primarily to boys or girls. To the extent that children are more likely to imitate those that are like them [@bussey1999social], finding that books containing boy and girl stereotypes are read to boys and girls, respectively, would suggest that gender stereotypes in books may present particularly potent opportunities for learning.


```{r child = "sections/00_corpus_desc_writeup.Rmd"}
```

```{r child = "sections/01a_judgment_word.Rmd"}
```

```{r child = "sections/01b_judgment_books.Rmd"}
```

```{r child = "sections/01c_judgment_validation.Rmd"}
```

```{r child = "sections/02_cooccurrence_stats.Rmd"}
```

```{r child = "sections/03_book_audience_writeup.Rmd"}
```

# General Discussion


What gender messages are conveyed by popular children’s books and who is being exposed to them? We constructed a corpus of 247 contemporary children’s books and analyzed the extent to which the books contain gender stereotypes. Using adult judgments of individual words, we found that over half of the words in the corpus tended to be associated with a particular gender. We then used word embedding models to explore the semantic associates of words in the corpus, finding that gender-biased words formed gender stereotypical categories (e.g., social interaction for females; physical interaction for males). Further, word gender biases elicited from adult judgments and more specific gender stereotypes (e.g., boys are relatively better at math, and girls are relatively better at reading) were reflected in the language statistics of the corpus itself, _and_ were more exaggerated than in adult fiction. At the book level, we found that books varied in their genderedness, and contained statistical regularities reflecting gender stereotypes (e.g., girl characters tended to do stereotypically girl activities). These statistical regularities were stronger for female stereotypes relative to male stereotypes. Finally, we derived a novel metric for measuring the gender distribution of a book’s audience using automated analysis of book reviews. Children tended to be exposed to books that conveyed gender stereotypes about their own gender. Our work provides the first quantitative assessment of the nature of gender messages within contemporary children’s books, and reveals that they contain many statistical regularities that could inform children’s understanding of gender stereotypes. 

There are several reasons to think that the statistical regularities we identified in children’s books may be shaping children’s gender stereotypes. First, many of the stereotypical patterns that we report are implicit in the distributional statistics of the text, rather than conveyed via explicit statements (“boys are better at math than girls”). The implicit nature of these messages may make them particularly difficult for adult readers to track or explicitly contradict. Second, children are exposed to books with a caregiver (compared to, e.g., watching TV). The caregiver's presence may signal implicit endorsement of these stereotypes as correct or desirable and lead the child to make to stronger inferences [@lewis2016understanding; @xu2007b]. Third, our data suggest that children tend to be exposed to books that contain gender stereotypes of their own gender presenting children with more information about own-gender-consistent associations. This may make gender-inconsistent preferences less familiar to children and therefore more difficult to emulate [@bussey1999social]. Filtered through children’s cognitive and social biases, children's books may therefore be a potent means of teaching children about gender stereotypes. 

Our work characterizes the messages in the text of children’s books and begins to address the role they play in socialization, but there are a number of open questions about the causal link between the statistical regularities we observe and the gender stereotypes that children form. Importantly, little is known about how children themselves perceive the messages contained within these books. In the work presented here, we primarily measure word gender bias via adult judgments, yet children do not have the extensive knowledge and experience that underlies adult judgments. The fact that word embedding models trained exclusively on the statistics of the children’s book corpus reflect adult-like word gender biases suggests that the adult gender biases could in principle be learned from sources like children’s book text, but it is an open question whether they actually do. Future work could more directly address these questions by eliciting child ratings of word gender, and by experimentally manipulating the statistics of children’s linguistic input about gender. 

An unanswered question from our data is whether the tendency for children to be read books matching their own gender is due to caregiver or child preferences. This question is important in light of recent data on gender development in transgender children [@gulgoz2019similarity]. Transgender children show strong identity with the gender they feel they are by three years of age. If transgender children play an active role in their own socialization [@martin2004children], our data suggest that children's books could be an early source of gender information for transgender children.

There is no doubt that shared reading has numerous benefits. However, our data show that embedded within contemporary children’s books are pervasive gender stereotypes --- indeed stronger than those found in adult-directed literature. Exposure to these language-embedded biases may lead to beliefs that help entrench gender disparities. Our work suggests caregivers may be able to influence children's  gender stereotype development through their choices in which books to read to their children. 



\newpage

# References

---
nocite: | 
 @time_100_kidbooks,@cvencek2011measuring, @skowronski2001comparative, @greenwald2002unified, @rudman2004gender, @nosek2002harvesting
...
 
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
