---
title             : "What might books be teaching young children about gender?"
shorttitle        : "Gender in children's books"

author:  
  - name          : "XXX"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "XXX"
    email         : "XXX"
  - name          : "XX"
    affiliation   : "2"
    email         : "cooperborken@wisc.edu"
  - name          : "XXX"
    affiliation   : "2"
    email         : "econverse@wisc.edu"
  - name          : "XXX"
    affiliation   : "2"
    email         : "glupyan@wisc.edu"
  - name          : "XXX"
    affiliation   : "2"
    email         : "seidenberg@wisc.edu "
affiliation:
  - id            : "1"
    institution   : "XXX"
  - id            : "2"
    institution   : "XXX"
    
author_note: | 
abstract: |
  We investigate how gender is represented in children’s books using a novel 200,000 word corpus comprising 247 popular, contemporary books for young children. Using human judgments and word co-occurrence data, we quantified gender biases of words in individual books and in the whole corpus. We find that children’s books contain many words that adults judge as gendered. Semantic analyses based on co-occurrence data yielded word clusters related to gender stereotypes (e.g., feminine: emotions; masculine: tools). Co-occurrence data also indicate that many books instantiate gender stereotypes identified in other research (e.g., girls are better at reading and boys at math). Finally, we used large-scale data to estimate the gender distribution of the audience for individual books, and find that children tend to be exposed to gender stereotypes for their own gender. Together the data suggest that children’s books may be an early source of gender associations and stereotypes.  
  
  STATMENT OF RELEVANCE:  Beliefs about gender, including stereotypes such as girls are better at reading and boys are better at math, originate in early childhood. Shared reading is an important source of information about about language and the world. It is therefore important to understand how gender is represented in books for young children (0-5 years old). The results from multiple analyses of a large set of popular books indicate that they are a rich source of information about gender, and that many express gender stereotypes, more strongly than adult fiction. These findings suggest that popular children’s books may be an underrecognized, inadvertent vehicle for perpetuating gender stereotypes and other gendered associations. 

  
keywords          : "reading, gender, language development"
wordcount         : "2060 (excluding methods and results but INCLUDING footnotes)"
bibliography      : ["references.bib"]
header-includes:
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{graphicx}
#  - \usepackage{booktabs}
  - \usepackage{tabu}
  - \usepackage{makecell}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \usepackage{pbox}
  - \usepackage{hyphsubst}
  - \floatplacement{figure}{t!} # make every figure with caption = t
  - \raggedbottom
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\textfraction}{.1}
  - \renewcommand{\floatpagefraction}{1}


figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
documentclass     : "apa6"
lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
#papaja::apa6_docx
  

---

```{r load_utility_packages, include = FALSE}
library(papaja)
library(rmarkdown)
library(tidyverse) 
library(here)
library(broom)
library(janitor)
library(kableExtra)
library(tidyboot)
library(ggrepel)
library(ggpmisc)
library(langcog)
library(glue)
library(esc)
library(compute.es)
library(broom.mixed)
library(numform)
library(rstatix)
library(lme4)
library(english)
source(here("writeups/paper/paper_helpers.R"))
```

```{r global_options, include = FALSE}
#redoc::redoc(highlight_outputs = FALSE)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      cache = FALSE,
                      echo = FALSE,
                      out.extra = "",
                      fig.pos = "t!",  sanitize = TRUE)
options(knitr.table.format = "latex")
```

Beliefs about gender-related characteristics develop early in childhood. By 24 months (girls) or 31 months (boys) children already exhibit knowledge of behaviors that are stereotypically feminine (e.g., vacuuming), masculine (e.g., building), and neutral [e.g., sleeping; @poulin2002men]. By age three, children distinguish individuals by gender, race, and age [@shutts2010social]. By age five, children have developed assumptions about gender “a constellation of stereotypes about gender (often amusing and incorrect) that they apply to themselves and others” [@martin2004children]. For example, preschoolers act in accordance with the stereotype that girls are better at reading while boys are better at math [@cvencek2011math], and that girls are less likely than boys to be “very, very smart” [@bian2017gender].

The sources of this knowledge are less well understood. Children’s interactions with adults and their observations of adult interactions are one [@hilliard2010differing]. Toys and activities are often gender stereotyped in home, daycare, and preschool social settings [@weisgram2014pink]. Gendered information is also conveyed via language. Children commonly receive verbal feedback from adults about gender-normative activities [e.g., girls more often about appearance and helping behaviors, boys about their size and physical skills; @chick2002impact]. Children are also sensitive to seemingly small differences in gender-related language [e.g., @chestnut2018girls; @moty2019unintended]. For example, @cimpian2011generic  found that when a novel game was introduced to children using a generic subject (“Girls are really good at a game called ‘gorp’”) they were more likely to associate it with the gender than when it was introduced with a specific subject (“There is a girl who is good at...”).

We examined a potentially rich yet underrecognized source of information about gender: children’s books. Reading to children (also called “shared reading”) has been widely encouraged because of its numerous benefits  [@bus1995joint; @duursma2008reading; @high2014literacy]. Shared reading marks the child’s entré to literacy and facilitates its development [@snow1998preventing]. It also promotes learning about aspects of language and the world beyond a child’s immediate experience [@mol2011read; @dickinson2012reading].  Reading with children could therefore be an important potential source of beliefs about gender.

Much previous work on how gender is represented in books has used “content analysis” methods that emphasize detailed analyses of a small number of texts. For example, @diekman2004learning examined 20 books for middle-schoolers categorized as “sexist” or “nonsexist”. College students each answered a 72-item questionnaire about one book. Questions probed whether books conveyed gender stereotypes and inequalities such as “Males, but not females, are shown as dominant” and “The book depicts female characters as the natural servants of male characters.” The results suggested that gender differences and inequalities were expressed even in books intended to be nonsexist.

Our goal was to conduct a broader analysis of gender representation in a large sample of common books for young (0-5 year old) children and to gain evidence about exposure to books by gender. We focused on the extent to which words in texts are associated with males vs. females, which we term the words’ “gender bias.” Some of these gender biases reflect well-known stereotypes,  for example  “pretty” (female) or “large” (male). By using both behavioral data and automated analyses of text characteristics, our approach provides a scalable and reproducible method of estimating gender bias without requiring explicit judgments of pre-specified properties of texts (as in studies such as Diekman & Murnen, 2004).

We begin by describing the construction and properties of the Wisconsin Children’s Book Corpus (WCBC). We then quantify gender biases in individual books and the corpus as a whole using two methods. Study 1 employed adult word-genderedness judgments, Study 2 statistical co-occurrences of words. The results indicate that books vary widely in degree of gender bias, ranging from strongly male to strongly female. Study 3 used analyses of gender biases in book reviews to estimate whether the books are being read primarily to boys or girls. Finding that books exhibiting male vs. female biases are more often read to boys and girls, respectively, would suggest that books may offer extensive as well as different opportunities for learning about gender. 



```{r child = "sections/00_corpus_desc_writeup.Rmd"}
```

```{r child = "sections/01a_judgment_word.Rmd"}
```

```{r child = "sections/01b_judgment_books.Rmd"}
```

```{r child = "sections/01c_judgment_validation.Rmd"}
```

```{r child = "sections/02_cooccurrence_stats.Rmd"}
```

```{r child = "sections/03_book_audience_writeup.Rmd"}
```

# General Discussion

What gender messages are conveyed by popular children’s books and who is being exposed to them? We constructed a corpus of 247 contemporary children’s books and analyzed the extent to which the books contain biased gender associations. Using adult judgments of individual words, we found that over half of the words in the corpus tended to be associated with a particular gender, and tended to cohere in gender stereotypical categories (e.g., social interaction for females; physical interaction for males). At the book level, we found that books varied in their gender associations, and that the associations tended to reflect gender stereotypes (e.g., girl characters tended to do stereotypically girl activities). Further, the language statistics of the corpus itself paralleled word gender biases seen in adult judgments and specific gender stereotypes (e.g., boys are better at math, and girls are better at reading). These biases were more exaggerated in the children’s book corpus, relative to adult fiction. Finally, we derived a novel metric for measuring the gender distribution of a book’s audience using automated analysis of book reviews. Children tended to be exposed to books that conveyed gender stereotypes about their own gender. Our work provides the first quantitative assessment of how gender is represented in contemporary children’s books and reveals that they contain many statistical regularities that could inform children’s understanding of gender stereotypes.

There are several reasons to think that the statistical regularities we identified in children’s books may be shaping children’s gender stereotypes. First, many of the stereotypical patterns that we report are implicit in the distributional statistics of the text, rather than conveyed via explicit statements (“boys are better at math than girls”). The implicit nature of these messages may make them particularly difficult for adult readers to track or explicitly contradict. Second, children are exposed to books with a caregiver (compared to, e.g., watching TV). The caregiver’s presence may signal implicit endorsement of these stereotypes as correct or desirable and lead the child to make stronger inferences [@lewis2016understanding; @xu2007b]. Third, our data suggest that children tend to be exposed to books that contain gender stereotypes of their own gender presenting children with more information about own-gender-consistent associations. This may make gender-inconsistent preferences less familiar to children and therefore more difficult to emulate [@bussey1999social]. Filtered through children’s cognitive and social biases, children’s books may therefore be a potent means of teaching children about gender stereotypes.

Our work characterizes the messages in the text of children’s books and begins to address the role they play in socialization, but there are a number of open questions about the causal link between the statistical regularities we observe and the gender stereotypes that children form. Importantly, little is known about how children themselves perceive the messages contained within these books. In the work presented here, we primarily measure word gender bias via adult judgments, yet children do not have the extensive knowledge and experience that underlies adult judgments. The fact that word embedding models trained exclusively on the statistics of the children’s book corpus reflect adult-like word gender biases suggests that the adult gender biases could in principle be learned from sources like children’s book text, but whether they are remains an open question. Future work could more directly address these questions by eliciting child ratings of word gender, and by experimentally manipulating the statistics of children’s linguistic input about gender.

One unanswered question from our data is how children learn stereotypes about other genders, given that they are largely read storybooks containing stereotypes aligning with their own gender. One possibility is that children are exposed to information about other genders from other sources, such as other kinds of media and direct interactions. Alternatively, children may in fact receive more information about their own gender, relative to other genders, and consequently have less precise intuitions about stereotypes related to other genders. It is also an open question whether the tendency for children to be read books matching their own gender is due to caregiver or child preferences. This question is important in light of recent data on gender development in transgender children [@gulgoz2019similarity]. Transgender children show strong identity with the gender they feel they are by age three. If transgender children play an active role in their own socialization  [@martin2004children], our data suggest that children’s books could be an early source of gender information for transgender children.

There is no doubt that shared reading has numerous benefits. However, our data show that contemporary children’s books also convey systematic information about gender, often (though not always) instantiating gender stereotypes — indeed some more strongly than in adult-directed literature. Caregivers may inadvertently promote the development of gender stereotypes via shared reading of books. Exposure to these language-embedded biases may lead to beliefs that help entrench gender biases and disparities. However, the variability of gender biases across books also suggests that caregivers may be able to influence children’s development of beliefs about gender through choice of books, an important issue for future research.


\newpage

# References

---
nocite: | 
 @time_100_kidbooks,@cvencek2011measuring, @skowronski2001comparative, @greenwald2002unified, @rudman2004gender, @nosek2002harvesting, @wagner2017factors
...
 
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
