---
title: Math words in kidbook corpus
subtitle: gender analysis
author: Molly Lewis 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: no
    toc: yes
---
  
******

```{r setup, include = F}
# load packages
library(knitr)
library(rmarkdown)
library(tidyverse)
library(googlesheets4)
library(broom)
library(langcog)
library(here)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, tidy = F, cache = F)
theme_set(theme_classic())

```

These analses are based on a list of 65 children's "mathy" words I generated in an ad hoc way. You can see them in a [google sheet here](https://docs.google.com/spreadsheets/d/1GTdLmxLhIEa4hWCGf_relV54CcuVrqORKE1dS3Ajy0Q/edit#gid=0). 

```{r}
MATH_WORDS <-  "1GTdLmxLhIEa4hWCGf_relV54CcuVrqORKE1dS3Ajy0Q"
math_words <- read_sheet(MATH_WORDS)
```

```{r}
CORPUS_PATH <- here("data/processed/books/tidy_full_corpus_all.csv")
corpus <- read_csv(CORPUS_PATH) %>%
  select(-corpus_type)
```

```{r}
REVIEWS_DATA_PATH <- here("data/processed/other/amazon_gender_scores.csv")
review_data <- read_csv(REVIEWS_DATA_PATH) 

by_book_review_data <- review_data %>%
  group_by(book_id, n_reviews_total, n_reviews_gendered,
           prop_review_gendered)  %>%
  summarize(addressee_gender_score_token =
              sum(n_female_token)/(sum(n_female_token) +
                                     sum(n_male_token))) %>%
  ungroup() %>%
  select(book_id, addressee_gender_score_token)
```


## Proportion math words predicted by audience gender

Each point in these analyses is a book. The audience gender measure is the one described in the paper, derived from Amazon Reviews. 

```{r}
total_spatial_words_by_type_types <- corpus %>%
  distinct(book_id, word) %>%
  left_join(math_words) %>%
  group_by(book_id, type) %>%
  summarize(n_mathy = n()) %>%
  ungroup() %>%
  complete(book_id, type, fill = list(n_mathy = 0))  

all_word_counts_types<- total_spatial_words_by_type_types %>%
  group_by(book_id) %>%
  summarize(total_words = sum(n_mathy, na.rm = T))

spatial_words_by_type_types <- total_spatial_words_by_type_types %>%
  filter(!is.na(type)) %>%
  left_join(all_word_counts_types) %>%
  mutate(prop_mathy = n_mathy/total_words,
         log_prop_mathy = log(prop_mathy),
         log_prop_mathy_corrected = log(prop_mathy + .001)) %>%
  left_join(by_book_review_data)

spatial_words_all_types <- spatial_words_by_type_types %>%
  group_by(book_id) %>%
  summarize(n_mathy = sum(n_mathy),
            total_words = total_words[1],
            addressee_gender_score_token = addressee_gender_score_token[1]) %>%
    mutate(prop_mathy = n_mathy/total_words,
           log_prop_mathy = log(prop_mathy),
           log_prop_mathy_corrected = log(prop_mathy + .001))
```

### All words
```{r}
ggplot(spatial_words_all_types, aes(x = addressee_gender_score_token, y = log_prop_mathy)) +
  geom_point(alpha = .8) +
  geom_smooth(method = "lm") +
  ylab("Log proportion math words") +
  xlab("Prop. female book audience")

spatial_words_all_types %>%
  filter(is.finite(log_prop_mathy)) %>%
  nest() %>%
  mutate(temp = map(data, ~tidy(cor.test(.$addressee_gender_score_token,
                                         .$log_prop_mathy)))) %>%
  select(-data) %>%
  unnest() %>%
  kable()
```

### Words by math type
```{r}
  
ggplot(spatial_words_by_type_types, aes(x = addressee_gender_score_token, y = log_prop_mathy)) +
  facet_wrap(~type, scales= "free_y") +
  geom_point(alpha = .8) +
  geom_smooth(method = "lm") +
  ylab("Log proportion math words") +
  xlab("Prop. female book audience")

spatial_words_by_type_types %>%
  filter(is.finite(log_prop_mathy)) %>%
  nest(-type) %>%
  mutate(temp = map(data, ~tidy(cor.test(.$addressee_gender_score_token,
                                         .$log_prop_mathy)))) %>%
  select(-data) %>%
  unnest() %>%
  kable()
```

## Proportion math words predicted by main character gender
```{r}
CHARACTER_DATA <- here("data/raw/other/character_gender_by_book.csv")
character_data <- read_csv(CHARACTER_DATA)

tidy_character_data <- character_data %>%
  select(book_id, char_main_gender) %>%
  mutate(char_main_gender = fct_recode(char_main_gender, 
                                          female = "F",
                                          male = "M",
                                          "indeterminate" = "AND",
                                          mixed = "MIXED")) 
```

### All words
```{r}
all_words_by_book_character <- spatial_words_all_types %>%  
  filter(is.finite(log_prop_mathy)) %>%
  left_join(tidy_character_data) %>%
  group_by(char_main_gender) %>%
  multi_boot_standard(col = "log_prop_mathy")

ggplot(all_words_by_book_character, aes(x = char_main_gender, y = mean)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Log proportion math words") +
  xlab("Main character gender")
```

### Words by math type
```{r}
words_type_by_book_character <- spatial_words_by_type_types %>%  
  filter(is.finite(log_prop_mathy)) %>%
  left_join(tidy_character_data) %>%
  group_by(char_main_gender, type) %>%
  multi_boot_standard(col = "log_prop_mathy")
  
ggplot(words_type_by_book_character, aes(x = char_main_gender, y = mean)) +
  facet_wrap(~type, scales= "free_y") +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Log proportion math words") +
  xlab("Main character gender") +
  theme(axis.text.x = element_text(angle = 90))
```
