
```{r}
library(tidyverse)
library(here)
library(numform)
library(rstatix)
library(lme4)
```

```{r bookmeans}
BOOK_MEANS_PATH <- here("data/processed/books/gender_token_type_by_book.csv")
gender_rating_by_book_mean_only <- read_csv(BOOK_MEANS_PATH)
overall_token_mean <- mean(gender_rating_by_book_mean_only$token_gender_mean)

CHARACTER_PATH <- here("data/raw/other/character_gender_by_book.csv")
characters <- read_csv(CHARACTER_PATH)  

book_means <- gender_rating_by_book_mean_only %>%
  left_join(characters %>% select(book_id,char_main_gender, 
                                  char_second_gender)) %>%
  filter(!(book_id %in% c("L105", "L112"))) # "Journey" and "Anno's Journey" are pictures books
```

## Study 1b: Gender bias in books
We next use the word gender bias judgments derived in Study 1a to quantify the genderedness of individual books. 

### Method and Results

```{r}
prop_present <- book_means %>%
  group_by(corpus_type) %>%
  multi_boot_standard(col = "prop_present_type") %>%
  filter(corpus_type == "all")

overall_gender_mean_token <- book_means %>%
  group_by(corpus_type) %>%
  multi_boot_standard(col = "token_gender_mean") %>%
  filter(corpus_type == "all")

overall_sd <- book_means %>% 
  filter(corpus_type == "all") %>%
  summarize(overall_sd = sd(token_gender_mean)) %>%
  pull(overall_sd) %>%
  round(2)

# forest plot below
all_book_means <- book_means %>%
  filter(corpus_type == "all") %>%
  mutate(char_main_gender = case_when(is.na(char_main_gender) ~ "none",
                                       TRUE ~ char_main_gender),
          char_main_gender = fct_recode(char_main_gender, 
                                          female = "F",
                                          male = "M",
                                          "indeterminate" = "AND",
                                          mixed = "MIXED"),
          char_main_gender = fct_relevel(char_main_gender,
                                              "female", "male", "indeterminate", "mixed", "none"))  %>%
  arrange(token_gender_mean) %>%
  mutate(gender_order = 1:n(), 
        gender_group = case_when(gender_order <= 20 ~ "male-biased",
                                 gender_order >= 109 &  gender_order <= 128 ~ "neutral",
                                 gender_order >= 230 ~ "female-biased"),
        gender_group = fct_relevel(gender_group, "female-biased", "neutral"),
        title = str_to_title(title))%>%
  filter(!is.na(gender_group))

```


Using estimates of word gender bias from adult judgments in Study 1, we calculated  overall gender bias score for each book as the mean gender bias score of all the words (tokens) it contained. On average, there were gender norms for `r pull(prop_present, mean)` ([`r pull(prop_present, ci_lower)`, `r pull(prop_present, ci_upper)`]) of all tokens in the books (see SI for details and additional analyses). The average gender score did not exhibit a strong bias (_M_ =  `r pull(overall_gender_mean_token, mean)` [`r pull(overall_gender_mean_token, ci_lower)`, `r pull(overall_gender_mean_token, ci_upper)`]), but there was substantial variability (_SD_ = `r overall_sd`): some books contained many more “masculine” words, other books contained many more “feminine” words.

```{r}
gender_by_corpus_type <- book_means %>%
  group_by(corpus_type) %>%
  multi_boot_standard(col = "token_gender_mean") %>%
  mutate_if(is.numeric, round, 2) %>%
  mutate(print_mean = glue("_M_ = {mean} [{ci_lower}, {ci_upper}]"))

corr_df <- book_means %>% 
  select(book_id, corpus_type,token_gender_mean) %>%
  spread(corpus_type, token_gender_mean, -1, fill = NA) 

all_char_corr <- get_tidy_corr_text(corr_df, "all", "char_only")
all_no_char_corr <- get_tidy_corr_text(corr_df, "all", "no_char")
char_no_char_corr <- get_tidy_corr_text(corr_df, "char_only", "no_char")
```



\autoref{fig:bookforest} includes data from a subset of books, the 20 with the highest feminine bias scores, the 20 with the highest masculine bias scores, and 20 from the neutral range. Data for all books are available in the SI. Measured in this way, the books clearly vary in genderedness, falling along a continuum. Books at the feminine end include *Olivia*, *Brave Irene*, and *Amelia Bedelia*; the masculine end includes *Dear Zoo*, *Curious George*, and *Good Dog, Carl*; neutrals include *In the Night Kitchen*, *Hippos Go Berserk*, and *Everyone Poops*. The feminine titles include more references to family members (mommy, sister, grandma); the masculine titles include more references to animals and non-familial characters. 

Differences in overall gender bias may be due to the distribution of content words without “intrinsic” gender (e.g., *beautiful*, *fight*) but also to differences in the occurrence of intrinsically gendered words such as names (*Jill*), pronouns (*her*), and relational/generic gender terms (e.g., *mom*, *lady*). We therefore calculated two additional bias measures, one including only the intrinsically-biased words (the character gender score) and the other including all words except the intrinsically-biased ones (the content gender score). Both the character (`r filter(gender_by_corpus_type, corpus_type == "char_only") %>% pull(print_mean)`; `r all_char_corr`) and content scores (`r filter(gender_by_corpus_type, corpus_type == "no_char") %>% pull(print_mean)`; `r all_no_char_corr`) were correlated with the overall gender score. Thus, both gendered content words and intrinsically gendered words contribute to the overall gender differences between books. 

Character and content scores had a moderate positive correlation with each other (`r char_no_char_corr`). Books with more feminine-biased (less masculine-biased) content words do tend to have more female names, pronouns and other intrinsically gendered words (\autoref{fig:charactercontentplot}). This finding suggests that gender biases reported by adults for content words are potentially inferable from the character associations of the content words in the book texts. 


```{r bookforest, fig.height = 8, fig.width = 8, fig.pos = "b", fig.cap = "Overall gender rating of  a subset of books, the 20 with the highest feminine bias scores, the 20 with the highest masculine bias scores, and 20 from the neutral range. Bias scores are calculated from the mean gender ratings of words in each book (tokens). The dashed line indicates the overall mean across books, and color indicates the gender of the primary character. Ranges are bootstrapped 95\\% CIs."}

ggplot(all_book_means, aes(x = reorder(title, token_gender_mean),
             y = token_gender_mean)) +
  geom_hline(aes(yintercept = mean(all_book_means$token_gender_mean)), linetype = 2) +
  facet_grid(gender_group ~ ., drop=TRUE,scales = "free_y") +
  coord_flip() +
  geom_linerange(aes(ymin = token_ci_lower, ymax = token_ci_upper), alpha = .3) +
  geom_point(aes(color = char_main_gender), size = 2) +
     scale_color_manual(
      values = c("pink"," lightblue", "lightgoldenrod1",  "palegreen", 
                      "lightgrey"), 
      name = "Primary character\ngender") +
  xlab("Book Title") +
  ylab("Book Gender Score (female-ness)") +
  ggtitle("Mean Gender Rating by Book") +
  theme_classic() +
  theme(axis.text.y = element_text(size = 5.5)) +
       # legend.position = c(.8,.2)) +
  guides(colour = guide_legend(override.aes = list(size=5)))
```


```{r charactercontentplot"}

LABELED_BOOKS <- c("L143", "M81", "M25", "M56",  "M44", "L142", "L239",
                  "L213", "M68", "M39", "M548", "L136", "L220",   "M49", "L111",  "L160")

book_labels <- gender_rating_by_book_mean_only %>% 
  distinct(book_id, title) %>% 
  filter(book_id %in% LABELED_BOOKS) %>%
  mutate(title = case_when(title == "GOODNIGHT, GOODNIGHT, CONSTRUCTION SITE" ~ 
                             "GOODNIGHT, GOODNIGHT,\nCONSTRUCTION SITE", TRUE~ title),
         title = case_when(title == "TRIANGLE (SHAPE TRILOGY)" ~ 
                             "TRIANGLE ", TRUE~ title),
           title = case_when(title == "CURIOUS GEORGE TAKES A JOB" ~ 
                             "CURIOUS GEORGE\nTAKES A JOB", TRUE~ title),
          title = case_when(title == "KATY AND THE BIG SNOW" ~ 
                             "KATY AND THE\nBIG SNOW", TRUE~ title),
         title = str_to_title(title))

charactercontentplot <- corr_df %>%
  left_join(book_labels) %>%
  filter(!is.na(char_only)) %>%
  mutate(point_class = case_when(!is.na(title) ~ "l", TRUE ~ "h")) %>%
  ggplot(aes(x = char_only, y = no_char)) +
    xlab("Character Gender Score") +
    ylab("Content Gender Score") +
      geom_smooth(method = "lm") +
  
    geom_point( aes( alpha = point_class)) +
    geom_text_repel(aes(label = title), size = 2.5, 
                  min.segment.length = 0) +
    scale_alpha_manual(values = c(.15, .9)) +
    ggtitle("Book Content vs.\nCharacter Gender Scores") +
    scale_x_continuous(breaks = c(1,2,3,4,5), 
                     label = c("1\n(male-\nbiased)",  "2" ,"3", "4", "5\n(female-\nbiased)")) +
    scale_y_continuous(breaks = c(2.7, 2.8,3.0,3.2,3.4), 
                     label = c("(male-\nbiased)",  "2.8", "3.0" ,"3.2", "3.4\n(female-\nbiased)"),
                     limits = c(2.68, 3.45) ) +
    annotate("text", x = 4, y = 2.75, label = paste0("italic(r) ==  ", 
                                                     f_num(cor(corr_df$no_char, corr_df$char_only, use = "complete"), 2)),
           color = "red", size = 4, parse = T) +
    theme_classic(base_size = 10) +
    theme(#axis.line = element_line(size = .8),
          legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
character_bar_data <- characters %>%
  select(book_id, char_main_singular, char_main_gender,
        char_second_singular, char_second_gender) %>%
  mutate_if(is.character, as.factor)   %>%
  mutate(char_second_gender2 = case_when(is.na(char_second_gender) ~ "none",
                                TRUE ~ as.character(char_second_gender)),
         char_main_gender2 = case_when(is.na(char_main_gender) ~ "none",
                                TRUE ~ as.character(char_main_gender)),
         char_second_gender2 = fct_recode(char_second_gender2, female = "F",
                                          male = "M",
                                          mixed = "MIXED",
                                          "indeterminate" = "AND"),
         char_main_gender2 = fct_recode(char_main_gender2, female = "F",
                                          male = "M",
                                          mixed = "MIXED",
                                          "indeterminate" = "AND"),
         char_second_gender2 = fct_relevel(char_second_gender2, "female",
                                          "male", "indeterminate", "mixed",
                                          "none"),
         char_main_gender2 = fct_relevel(char_main_gender2,
                                              "female", "male", "indeterminate",
                                         "mixed", "none"))

all_gendered_counts <- character_bar_data %>%
  count(char_main_gender2) %>%
  mutate(perc = paste0(round(n/sum(n) *100,0), "%"),
         perc_print = glue("{n} ({perc})"))

mf_gendered_counts <- data.frame(n_gendered = sum(all_gendered_counts %>% filter(char_main_gender2 == "female") %>% pull(n),
                                                  all_gendered_counts %>% filter(char_main_gender2 == "male") %>% pull(n))) %>%
  mutate(perc_gendered = paste0(round(n_gendered/sum(all_gendered_counts$n) * 100,0), "%"),
         print_total_gendered = glue("{n_gendered}/{sum(all_gendered_counts$n)}; {perc_gendered}")) %>%
  pull(print_total_gendered)
                      
primary <- character_bar_data %>%
  filter(char_main_gender %in% c("F", "M")) %>%
  count(char_main_gender)

primary_chi <- chisq.test(primary$n) %>% 
   tidy()
  
primary_chi_text <- primary_chi %>%
   mutate(p_char = round(p.value,2),
            x_equality_sign = case_when(p_char < .001 ~ ", _p_ < .001", 
                                     TRUE ~ paste0(", _p_ = ", p_char)),
          x_print_text = paste0("(", parameter, ") = ", as.character(round(statistic, 2)),
                                x_equality_sign, "")) %>%
  pull(x_print_text)

primary_chi_es_text <- esc_chisq(chisq = primary_chi$statistic, 
          totaln = sum(primary$n), es.type = "d") %>%
  as.data.frame() %>%
  mutate_if(is.numeric, round,2) %>%
  mutate(d_print_text = glue("_d_ = {es} [{ci.lo}, {ci.hi}]")) %>%
  pull(d_print_text)
```

Whereas the character gender score above reflects the extent to which males and females are mentioned in a book, the gender of the story protagonist may be particularly influential for children. For each book, we manually coded the name of the primary protagonist character(s) and the character’s gender as determined from the text (i.e., not based on the illustrations). A character was considered a protagonist if that character was the primary agent of the story, even if in a collaborative fashion with another protagonist. The main character(s) were classified as either female, male, mixed, or indeterminate [@wagner2017factors]. If there was more than one primary character, and their gender composition was heterogeneous, that group was classified as mixed. If a given primary character had a gender that could not be determined, no gender attribute was assigned (“indeterminate”). Two research assistants and the second author coded character gender.  Coders agreed on the protagonist type for 97% of books. Discrepancies were resolved through discussion.

```{r characterplot, out.width = "\\textwidth", fig.cap = "(a) Mean content gender score for each book as a function of mean character gender score. Higher values indicate a greater propensity for female-biased words. Error bar shows the standard error of the linear model fit. (b) Distribution across books of content gender score as a function of primary character gender. Colored points show individuals books (one point excluded for visibility). Dashed line shows content gender score grand mean. Black points and ranges show mean and bootstrapped 95\\% percent confidence intervals for books of each primary character gender type."}
violin_data <- book_means %>%
    mutate_if(is.character, as.factor)   %>%
    mutate(char_main_gender2 = case_when(is.na(char_main_gender) ~ "none",
                                TRUE ~ as.character(char_main_gender)),
           char_main_gender2 = fct_recode(char_main_gender2, female = "F",
                                          male = "M",
                                          mixed = "MIXED",
                                          "indeterminate" = "AND"),
          char_main_gender2 = fct_relevel(char_main_gender2, "male", "female", "indeterminate",
                                         "mixed", "none"))  %>%
  filter(corpus_type == "no_char") 

content_means_for_violin <- violin_data %>%
  group_by(char_main_gender2) %>%
  multi_boot_standard(col = "token_gender_mean")

violin_plot1 <- ggplot(violin_data %>% filter(token_gender_mean < 3.5), # exclude one point
         aes(x = char_main_gender2, y = token_gender_mean, color = fct_rev(char_main_gender2))) +
  geom_violin() +
  geom_jitter(alpha = .7, size = 1) +
  geom_point(data = content_means_for_violin, 
                  aes(y = mean), 
                  color = "black",
             size = 2) +
  geom_linerange(data = content_means_for_violin, 
                  aes(y = mean, ymin = ci_lower, ymax = ci_upper), 
                  color = "black", 
                 size = .6) +
   geom_hline(aes(yintercept = mean(violin_data$token_gender_mean)),
              linetype = 2, size = .4) +
   scale_color_manual(
      values = rev(c(" lightblue","pink", "lightgoldenrod1", 
                     "lightgreen", 
                      "lightgrey")), 
      name = "Main character\ngender", guide = F) +
    scale_y_continuous(breaks = c(2.7, 2.8,3.0,3.2,3.4), 
                     label = c("(male-\nbiased)",  "2.8", "3.0" ,"3.2", "3.4\n(female-\nbiased)"),
                     limits = c(2.68, 3.45) ) +
  ggtitle("Book Content vs.\nCharacter Gender") +
  xlab("Primary Character Gender") +
  ylab("Content Gender Score") +
  theme_classic(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cowplot::plot_grid(charactercontentplot, violin_plot1, ncol = 2,
                             labels = c('a','b'), label_size = 14, rel_widths = c(1.6,1))

```


```{r charbookmeans}
book_gender_descriptives <- book_means %>%
  filter(char_main_gender %in% c("M", "F"))  %>%
  group_by(char_main_gender, corpus_type) %>%
  multi_boot_standard(col = "token_gender_mean") %>%
  mutate_if(is.numeric, round, 2) %>%
  mutate(print_mean = glue("_M_ = {mean} [{ci_lower}, {ci_upper}]"))


# do character scores differ from mean for males? and females?
char_data <- book_means %>% 
  filter(corpus_type == "char_only")

char_t_test_F <- get_tidy_one_sample_t_test_text(
  filter(char_data, char_main_gender == "F") %>% 
                           pull(token_gender_mean),
  mean(pull(char_data, token_gender_mean)))

char_t_test_M <- get_tidy_one_sample_t_test_text(
  filter(char_data, char_main_gender == "M") %>% 
                           pull(token_gender_mean),
  mean(pull(char_data, token_gender_mean)))


# do content scores differ from mean for males? and females?
no_char_data <- book_means %>% 
  filter(corpus_type == "no_char")
  
content_t_test_F <- get_tidy_one_sample_t_test_text(
  filter(no_char_data, char_main_gender == "F") %>% 
                           pull(token_gender_mean),
  mean(pull(no_char_data, token_gender_mean)))

content_t_test_M <- get_tidy_one_sample_t_test_text(
  filter(no_char_data, char_main_gender == "M") %>% 
                           pull(token_gender_mean),
  mean(pull(no_char_data, token_gender_mean)))
```



About half of the books (`r mf_gendered_counts`) had gendered primary characters that were exclusively male or exclusively female. Two-thirds of these books had male primary characters (_N_ = `r all_gendered_counts %>% filter(char_main_gender2 == "male") %>% pull(n)`; $\chi^2$`r primary_chi`; `r primary_chi_es_text`). Of the remaining books, `r all_gendered_counts %>% filter(char_main_gender2 == "indeterminate") %>% pull(perc_print)` had main characters(s) of indeterminate gender, `r all_gendered_counts %>% filter(char_main_gender2 == "mixed") %>% pull(perc_print)` had main characters of mixed genders, and `r all_gendered_counts %>% filter(char_main_gender2 == "none") %>% pull(perc_print)` had no main character(s). We then examined book genderedness as a function of the gender of the primary character, using both content and character scores. Books with female primary characters tended to have higher female content scores (`r book_gender_descriptives %>% filter(char_main_gender == "F", corpus_type == "no_char") %>% pull(print_mean)`; `r content_t_test_F`), compared to the overall averages, whereas books with male primary characters  tended to have relatively higher male content scores (`r book_gender_descriptives %>% filter(char_main_gender == "M", corpus_type == "no_char") %>% pull(print_mean)`; `r content_t_test_M`; \autoref{fig:characterplots}a). Notably, however, there was a large degree of variability in content scores across books of a given gender (female: $SD$  = `r filter(char_data, char_main_gender == "F") %>% pull(token_gender_mean) %>% sd()`; male: $SD$  = `r filter(char_data, char_main_gender == "M") %>% pull(token_gender_mean) %>% sd()`), suggesting that while some books contain gender stereotypes, others do not. 


Our findings suggest that books vary appreciably along the dimension of gender in terms of both their content and characters. The gender distribution of characters we observe is comparable to that reported previously in a smaller sample of books (Wagner, 2017).  Together, the gender character and gender content data provide converging evidence that information about gender associates of content words is present in the text of children’s books: Books with female characters tend to have content stereotypically associated with females, whereas books with male characters tend to have content stereotypically associated with males. 

